Modular Model-Checking for Fault-Tolerant Distributed System Implementations

### 2016-11-25 doc/om-paper/paper.tex
paper reminders
added ben's benchmark results.
add dashes for no data.
reworked intro.

The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,csl-93-2,pvs}. If formal verification is to be introduced into the workflow of system designers, though, we need automated methods and programmatic verification of \emph{implementations}. (Mostly) automated proof techniques are required to reduce the need for specialized verification expertise. System designers create software and hardware implementations to test, simulate, and deploy. Discrepancies between implementations can formal models can easily arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific. Furthermore, as an implementations are tweaked and modified to explore the design space, it is easy for the formal models to fall out of synch.

To address this problem, we envision building a domain-specific language (DSL) for specifying and exploring distributed fault-tolerant system implementations. We envision a DSL that generates both code for simulation, testing, and deployment, as well as formal models for verification (we say more about the DSL in Section~\ref{sec:conclusions}).

The first step in realizing this vision, however, and the one we address in this paper, is to develop an intermediate representation (IR) for the DSL to target when generating formal models. Our target for the IR is one supported by infinite-state model-checking. With proof techniques like IC3~\cite{} and $k$-induction~\cite{}, model-checking scales further than ever before. We envision an IR that allows for specifying system behavior, assumptions, and requirements, but also includes invariants that can be systematically generated by the DSL. Finally, the IR must support easy refactoring and modifications to the design. Making small changes to system specification---for example, modifying the fault model---should require only localized changes to the IR representation, including proofs. In short, the IR should be modular.

In this work, we combine a number of ideas known in the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems. We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and more recently popularized by Rushby for model-checking; and (3) \emph{abstract state machines} for abstracting complex systems.

To demonstrate the model, we describe a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{}, implementing the Boyer-Moore Fast Majority Vote algorithm~\cite{}. The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked implementation of the classic algorithm. Then, we modify the algorithm along axes such as the fault model, topology, node behavior, and show that only local modifications are required to the model and proofs.

@@

\begin{figure}
  \centering
  \includegraphics[width=8cm]{figs/benchmark}
  \caption{Comparison of number of nodes vs. verification time.}
  \label{fig:compare}
\end{figure}

% ------------------------------------------------------------
\section{Toward an Architectural Domain-Specific Language}\label{sec:adsl}
Not sure if we'll have space for this. Maybe just move to the conclusions?

@@

>>>>

The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,csl-93-2,pvs}. If formal verification is to be introduced into the workflow of system designers, though, we need more automated methods. (Mostly) automated proof techniques are required to reduce the need for specialized verification expertise. We also need programmatic verification of \emph{implementations}. System designers create software and hardware implementations to test, simulate, and deploy. Discrepancies between implementations and algorithmic models can arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific. Furthermore, as an implementations are tweaked and modified to explore the design space, it is easy for the formal models to become inconsistent.

There are at least two classes of abstractions that separate traditional models of fault-tolerant distributed algorithms from their implementations. One is to intertwine the environmental model with the system description. For example, the behavior of nodes is naturally specified as a transition system in which transitions are guarded by the node's fault state. But faults are part of the environment; in an implementation, does not change behavior based on its faults. Another class of abstractions is used to simplify models. For example, message passing might be abstracted with shared state, or a node's local behavior---e.g., a majority vote---is elided and instead, the output is constrained by a \emph{specification} of a majority vote.

In this paper, we develop a formal model to address these concerns. We combine ideas from the literature and folklore to build scalable and modular formal models of fault-tolerant distributed systems. We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and popularized by Rushby for model-checking; and (3) \emph{abstract state machines}, useful for abstracting counterexamples. The model is suitable for infinite-state model-checking. With proof techniques like IC3 and $k$-induction, model-checking scales further than ever before~\cite{pdr-kind}.

To demonstrate the model in Section~\ref{sec:byz}, we present a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{}, implementing the Boyer-Moore Fast Majority Vote algorithm for algorithm~\cite{}. The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked \emph{implementation} of the algorithm. In Section~\ref{sec:experimental}, we give performance metrics for the model, and we modify the implementation by changing the fault model, topology, node behavior, and show that only local modifications are required to the model and proofs.

%% To address this problem, we envision building a domain-specific language (DSL) for specifying and exploring distributed fault-tolerant system implementations. We envision a DSL that generates both code for simulation, testing, and deployment, as well as formal models for verification. That way, the formal models verified and code executed stay in synch, and abstractions are systematically made. We say more about the DSL in Section~\ref{sec:conclusions}.

@@

\begin{figure}
  \centering
  \begin{tabular}{r||c|c|c|c|c|c|c|c|c|c|}
      \multicolumn{11}{c}{Receivers} \\
      \cline{2-11}
   Relays &   1 &   2 &   3 &   4 &   5 &   6 &   7 &   8 &   9 &   10 \\
      \hline \hline
      1   & 13  & 18  & 23  & 29  & 35  & 43  & 54  & 70  & 90  & 129 \\
      \hline
      2   & 17  & 26  & 38  & 52  & 68  & 89  & 115 & 159 & 227 & -   \\
      \hline
      3   & 21  & 40  & 61  & 85  & 120 & 170 & 246 & 410 & -   & -   \\
      \hline
      4   & 27  & 58  & 98  & 195 & 213 & 628 & 855 & -   & -   & -   \\
      \hline
      5   & 36  & 93  & 292 & 389 & 1940& TO  & -   & -   & -   & -   \\
      \hline
      6   & 56  & 303 & 976 & TO  & TO  & -   & -   & -   & -   & -   \\
      \hline
      7   & 114 & 1993& TO  & TO  & -   & -   & -   & -   & -   & -   \\
      \hline
      8   & 338 & TO  & TO  & -   & -   & -   & -   & -   & -   & -   \\
      \hline
      9   & 895 & TO  & -   & -   & -   & -   & -   & -   & -   & -   \\
      \hline
     10   & TO  & -   & -   & -   & -   & -   & -   & -   & -   & -   \\
      \hline
  \end{tabular}
  \label{fig:benchmark}
  \caption{Benchmark of full proof computation time for OMH(1) implementation. Times are in seconds with a timeout (TO) limit of one hour. Dashes ('-') denote no benchmark was run.}
\end{figure}

@@

\url{https://people.eecs.berkeley.edu/~apanda/assets/papers/pldi16.pdf}
